On the presentation of orthogonal groups over polynomial rings

### 2019-05-21 dl-paper.tex
Sketched the proof of Lemma on descent

\begin{rem} \label{Pstar-char} The above lemma allows us to characterize $P_\alpha(*)$ as follows: it consists of images under $j_+$ of those elements 
 $g \in \overline{\St}(\Phi, A[X], M[X])$ for which $ev_{X=0}^*(g)$ lies in the subgroup $C_\alpha(A, M) := \Img(\overline{\St}(Z_0(\alpha), A, M) \to \overline{\St}(\Phi, A, M)) \ltimes \hat{U}(Z_+(\alpha), M)$ 
 (i.\,e. the subgroup generated by elements $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$ 
 and $Z_+(\alpha)$, $Z_0(\alpha)$ are as in~\cite[Proposition~5.7]{St71}). \end{rem}

@@

The next step of the proof is to define for $\alpha \in \Phi$, $a \in A$ a partial function $T_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for those triples $(g, h, u)$ for which $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$, then $T_\alpha(aX^{-1})$ is given by the following identity:
\[ T_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\]
It is easy to verify that $p(T_\alpha(aX^{-1}) (g, h, u)) = S_\alpha(a, p(g, h, u)).$

\begin{lemma} $T_\alpha(aX^{-1})$ descends to a well-defined total function $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $c = [(g, h, u)] \in \overline{V}$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev_{X=0}(g_1))^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $(g, h \cdot i_+(h_0), u)$ is the representative of $c$ on which $T_\alpha(aX^{-1})$ is defined.

 Next, let us show that the value of $T_\alpha(aX^{-1})$ does not depend on the choice of a representative.
 Let $(g, h_1, u), (g, h_2, u)$ be a pair of representatives of the same $c \in \overline{V}$ for which both $p(g, h_1, u)$ and $p(g, h_2, u)$ lie in $P_\alpha(*)$.
 By definition, $h_1 h_2^{-1} = i_-(h_0)$, for $h_0 \in \overline{\St}(\Phi, A, M)$, moreover, $p(g, h_1, u) \cdot p(g, h_2, u)^{-1} = j_- i_-(h_0) = j_+ i_+(h_0) \in P_\alpha(*)$,
 Clearly, $ev_{X=1}(j_+ i_+ (h_0)) = ev_{X=1}(i_+(h_0)) = ev_{X=0}(i_+(h_0)) = h_0$.
 On the other hand, by~\cref{Pstar-char} there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $ev_{X=0}(g_1) \in C_\alpha(A, M)$
  and $j_+(g_1) = j_+ i_+ (h_0)$.
 Now we have: $h_0 = ev_{X=1}(j_+(g_1)) = ev_{X=1}(g_1)$

 %Invoking injective stability for $\K_1(\Phi, -)$ and surjective stability $\K_2(\Phi, -)$ we can find $h_0 \in \overline{\St}(Z_0(\alpha), A, M) \ltimes U(Z_+(\alpha), M)$ such that $\pi(h_0) = \pi(j_-(h_1 h_2^{-1}))$.
\end{proof}

>>>>

\begin{rem} \label{Pstar-char} The above lemma allows us to characterize $P_\alpha(*)$ as follows: it consists of images under $j_+$ of those elements 
 $g \in \overline{\St}(\Phi, A[X], M[X])$ for which $ev_{X=0}^*(g)$ lies in the subgroup $Z_\alpha(A, M) := \Img(\overline{\St}(Z_0(\alpha), A, M) \to \overline{\St}(\Phi, A, M)) \ltimes \hat{U}(Z_+(\alpha), M)$ 
 (i.\,e. the subgroup generated by elements $x_{\beta}(m),\ \beta \in Z_+(\alpha)$ and $z_{\gamma}(m, \zeta),\ \gamma \in Z_0(\alpha),$ where $m \in M$, $\zeta \in A$ 
 and $Z_+(\alpha)$, $Z_0(\alpha)$ are as in~\cite[Proposition~5.7]{St71}). \end{rem}

@@

The next step of the proof is to define for $\alpha \in \Phi$, $a \in A$ a partial function $t_\alpha(aX^{-1}) \colon V \not\to V$.
This function is defined for those triples $(g, h, u)$ for which $p(g, h, u) \in P_\alpha(*) \subseteq \overline{G}_M^{\geq 0}$.
If $p(g, h, u) \in P_\alpha(m)$ for some $m \in M$, then $t_\alpha(aX^{-1})$ is given by the following identity:
\[ t_\alpha(aX^{-1}) (g, h, u) = \left( x_\alpha(aX^{-1})\cdot g ,\ h \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right).\]
It is easy to verify that $p(t_\alpha(aX^{-1}) (g, h, u)) = S_\alpha(a, p(g, h, u)).$

\begin{lemma} Let $\Phi$ be any root system of rank $\geq 3$, $\alpha \in \Phi$ and $(A, M)$ be a local pair.
 Then the partial function $t_\alpha(aX^{-1}) \colon V \not \to V$ descends to a well-defined total function 
  $T_\alpha(aX^{-1}) \colon \overline{V} \to \overline{V}$. \end{lemma}
\begin{proof}
 First of all, let us show that the resulting function is total. 
 Fix $c = [g, h, u] \in \overline{V}$. Since $p(g, h, u) \in \overline{G}^{\geq 0}_M$ there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $j_+(g_1) = p(g, h, u).$
 Set $h_0 = ev^*_{X=0}(g_1)^{-1}$, then, clearly, $g_1 \cdot i_+(h_0) \in K(A[X], M[X])$ and 
 $p(g, h \cdot i_-(h_0), u) = j_+(g_1) \cdot j_-(i_-(h_0)) = j_+(g_1 \cdot i_+(h_0)) \in j_+(K(A[X], M[X])) \subseteq P_\alpha(*)$,
  which shows that $(g, h \cdot i_+(h_0), u)$ is the representative of $c$ on which $t_\alpha(aX^{-1})$ is defined.

 Next, let us show that the value of $T_\alpha(aX^{-1})$ does not depend on the choice of a representative.
 Let $(g, h_1, u), (g, h_2, u)$ be a pair of representatives of the same $c \in \overline{V}$ for which both $p(g, h_1, u)$ and $p(g, h_2, u)$ lie in $P_\alpha(*)$.
 By definition, $h_1^{-1} h_2 = i_-(h_0)$, for some $h_0 \in \overline{\St}(\Phi, A, M)$, moreover, 
  $p(g, h_1, u)^{-1} \cdot p(g, h_2, u) = j_- i_-(h_0) = j_+ i_+(h_0) \in P_\alpha(*)$.
 By~\cref{Pstar-char} there exists $g_1 \in \overline{\St}(\Phi, A[X], M[X])$ such that $g_0 := ev^*_{X=0}(g_1) \in Z_\alpha(A, M)$ and $j_+(g_1) = j_+ i_+ (h_0)$.
 From the last equality and the injectivity of the map $G(\Phi, A[X]) \to G(\Phi, A[X, X^{-1}])$ we obtain that
 the projections of $g_0$, $g_1$ and $h_0$ in $G(\Phi, A[X, X^{-1}])$ are equal, therefore $g_0 \cdot h_0^{-1} \in \overline{\K_2}(\Phi, A, M)$. 
 By~\cite[Theorem~2.5]{Ste73} the latter subgroup is generated by relative Steinberg symbols $\{ a, m \}_\beta$, where $\beta$ is any fixed long root of $\Phi$.
 It is not hard to check using e.\,g. the table on p.~986 of~\cite{St71} that $Z_0(\alpha)$ always contains a long root, therefore
  $\overline{\K_2}(\Phi, A, M)$ is contained in $Z_\alpha(A, M)$ by ??? hence $h_0 \in Z_\alpha(A, M)$.
 Thus, we have shown that $i_-(h_0)$ is centralized by $X_\alpha(A[X^{-1}])$ therefore
 \begin{multline} \nonumber
  T_\alpha(aX^{-1})[g, h_1, u] = \left[ x_\alpha(aX^{-1})\cdot g ,\ h_1 \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right] = \\
  = \left[ x_\alpha(aX^{-1})\cdot g ,\ h_1 i_-(h_0) \cdot x_\alpha\left(-\tfrac{aX^{-1}}{1 + am}\right),\ u \cdot (1 + am)\right] = T_\alpha(aX^{-1})[g, h_2, u]. \qedhere
 \end{multline} \end{proof}
